# GT; 18/07/2024;

---
title: "R Notebook"
output: html_notebook
---

################################################################################
##### Some preliminaries ;

```{r}
library(ggcorrplot)
library(ggeffects)
library(ggExtra)
library(ggpubr)
library(ggridges)
library(ggthemes)
library(lme4)
library(mlogit)
library(qwraps2)
library(sjPlot)
library(support.CEs)
library(survival)
library(texreg)
library(tidyverse)

# options ggplot2 ;
theme_set(theme_bw())

################################################################################
# load (recoded) data ;
rm(list=ls())
exp_i   <- read.csv("REC_RESP_NOL.csv", header=TRUE)
exp_ij  <- read.csv("REC_EXP_NOL_ij.csv", header=TRUE)
exp_ij_ <- read.csv("REC_EXP_ij.csv", header=TRUE)

# within-perspective ;
exp_risk<- read.csv("REC_EXP_RISK_ij.csv", header=TRUE)
exp_ra  <- read.csv("REC_EXP_RA_ij.csv", header=TRUE)

################################################################################
# split into experiments 1-5 ;
exp_1   <- filter(exp_ij, exp==1) 
exp_2   <- filter(exp_ij, exp==2) 
exp_3   <- filter(exp_ij, exp==3) 
exp_4   <- filter(exp_ij, exp==4) 
exp_5   <- filter(exp_ij, exp==5) 

# split experiment 5 into vignettes 1-4 ;
exp_5_1 <- filter(exp_5, vignette==1)
exp_5_2 <- filter(exp_5, vignette==2)
exp_5_3 <- filter(exp_5, vignette==3)
exp_5_4 <- filter(exp_5, vignette==4)
```


################################################################################
##### EXPERIMENT 1 (TWO PRECISE PARTIES)
```{r}
# EXP1: vote shares of p_A (0) and p_B(1) ;
table(exp_1$vote)
prop.table(table(exp_1$vote))
```

```{r}
# EXP1: share of spatial voters ;
table(exp_1$vote_spa)
prop.table(table(exp_1$vote_spa))
t.test(exp_1$vote_spa, mu=.5)
```

```{r}
# EXP1: bivariate model ;
m1  <- glm(vote ~ ut_B_A, data=exp_1, family = binomial(link = "logit"))
summary(m1)
gofm(m1)
```

```{r}
m1_sq  <- glm(vote ~ ut_B_A_sq, data=exp_1, family = binomial(link = "logit"))
summary(m1_sq)
gofm(m1_sq)
```
```{r}
# CORRECT ;
ta <- prop.table(table(true=exp_1$vote, pred = round(fitted(m1))))
correct <- ta[1,1] + ta[2,2]
ta 
correct
```

```{r}
# EXP1: marginal effects ; 
ggpredict(m1, "ut_B_A") %>% 
  plot(ci.style="ribbon", add.data=TRUE, jitter=0.02) + 
  xlab(expression(paste("utility difference ",u(v,p[B]),'-',u(v,p[A])))) + 
  ylab("vote probability [Pr(v=B)]") +
  ggtitle("") + theme_bw()
```
################################################################################
##### EXPERIMENT 2 (A Precise and a Risky Party)
```{r}
# EXP2: vote shares of p(0) and p'_B' (1) ;
table(exp_2$vote_unc)
prop.table(table(exp_2$vote_unc))
t.test(exp_2$vote_unc, mu=.5)
```

```{r}
# EXP2: share of spatial voters ;
table(exp_2$vote_spa)
prop.table(table(exp_2$vote_spa))
t.test(exp_2$vote_spa, mu=.5)
```
```{r}
# EXP2: comparing E1 and E2; only identical positions ;
table(exp_2$switch_1_2)
prop.table(table(exp_2$switch_1_2))
t.test(exp_2$switch_1_2, mu=.5)
t.test(ifelse(exp_2$switch_1_2==0,1,0), mu=.5)
```

```{r}
# EXP2: bivariate model with p and p' ;
m2 <- glm(vote_unc ~ ut_unc_cer, data=exp_2, family = binomial(link = "logit"))
summary(m2)
gofm(m2)
```

```{r}
# EXP2: bivariate model with p and p' (squared);
m2_sq <- glm(vote_unc ~ ut_unc_cer_sq, data=exp_2, family = binomial(link = "logit"))
summary(m2_sq)
gofm(m2_sq)
```


```{r}
# EXP2: marginal effects  ;
ggpredict(m2, "ut_unc_cer") %>% 
  plot(ci.style="ribbon", add.data=TRUE, jitter=0.02) + 
  xlab("utility difference u(c[p'])-u(p)") + 
  ylab("vote probability [Pr(v=p')]") +
  ggtitle("") + theme_bw()

ggsave("../LaTeX/_fig_m2.pdf", plot=last_plot(), width = 12, height = 12, units = "cm")
```

```{r}
# EXP2: multivariate model (with interaction) ;
m2_pl  <- glm(vote_unc ~ ut_unc_cer, 
              data=exp_2, family = binomial(link = "logit"))
m2_add <- glm(vote_unc ~ ut_unc_cer + (lr_i_unc + risk + education), 
              data=exp_2, family = binomial(link = "logit"))
m2_int <- glm(vote_unc ~ ut_unc_cer * (lr_i_unc + risk + education), 
              data=exp_2, family = binomial(link = "logit"))

screenreg(list(m2_pl, m2_add, m2_int), digits=3, dcolumn = TRUE, booktabs = TRUE, 
       include.bic=FALSE, use.packages = FALSE, single.row=TRUE, caption.above=TRUE, 
       label = "TAB_MODELS", 
       caption = "Determinants of Choice Under Risk", 
       custom.coef.names = c("Intercept", "$u(p')-u(p)$", "uncertainty", "risk", "education", "$[u(p')-u(p)]$ * uncertainty", "$[u(p')-u(p)]$ * risk", "$[u(p')-u(p)]$ * education"),
       custom.model.names=c("Model 1 (bi-variate)", "Model 2 (additive)", "Model 3 (interactive)"),
       custom.note="Notes: Coefficients are obtained from a series of bi- and multivariate logit models based on E2.") 
```

```{r}
# EXP2: multivariate model (with interaction, squared) ;
m2_pl_sq  <- glm(vote_unc ~ ut_unc_cer, 
              data=exp_2, family = binomial(link = "logit"))
m2_add_sq <- glm(vote_unc ~ ut_unc_cer + (lr_i_unc + risk + education), 
              data=exp_2, family = binomial(link = "logit"))
m2_int_sq <- glm(vote_unc ~ ut_unc_cer * (lr_i_unc + risk + education), 
              data=exp_2, family = binomial(link = "logit"))

screenreg(list(m2_pl_sq, m2_add_sq, m2_int_sq), digits=3, dcolumn = TRUE, booktabs = TRUE, include.bic=FALSE, use.packages = FALSE, single.row=TRUE, caption.above=TRUE, 
       label = "TAB_MODELS", 
       caption = "Determinants of Choice Under Risk", 
       custom.coef.names = c("Intercept", "$u(p')-u(p)$", "uncertainty", "risk", "education", "$[u(p')-u(p)]$ * uncertainty", "$[u(p')-u(p)]$ * risk", "$[u(p')-u(p)]$ * education"),
       custom.model.names=c("Model 1 (bi-variate)", "Model 2 (additive)", "Model 3 (interactive)"),
       custom.note="Notes: Coefficients are obtained from a series of bi- and multivariate logit models based on E2.") 
```

```{r}
# EXP2: interaction with lr_i_unc ;
fig_m2_unc <- ggpredict(m2_int, c("ut_unc_cer", "lr_i_unc")) %>% 
  plot(colors="bw", add.data=TRUE, jitter=0.02) + 
  xlab("u(c[p'])-u(p)") + 
  ylab("vote probability [Pr(v=p')]") +
  ggtitle("Uncertainty") +
  theme(legend.position="bottom", legend.title=element_blank())
fig_m2_unc
```

```{r}
# EXP2: interaction with risk ;
fig_m2_risk <- ggpredict(m2_int, c("ut_unc_cer", "risk")) %>% 
  plot(colors="bw", add.data=TRUE, jitter=0.02) + 
  xlab("u(c[p'])-u(p)") +  
  ylab("") +
  ggtitle("Risk") + 
  theme(legend.position="bottom", legend.title=element_blank())
fig_m2_risk
```

```{r}
# EXP2: interaction with education ;
fig_m2_educ <- ggpredict(m2_int, c("ut_unc_cer", "education")) %>% 
  plot(colors="bw", add.data=TRUE, jitter=0.02) + 
  xlab("u(c[p'])-u(p)") +  
  ylab("") +
  ggtitle("Education") + 
  theme(legend.position="bottom", legend.title=element_blank())
fig_m2_educ
```

```{r}
# EXP2: combined marginal effects plots ;
ggarrange(fig_m2_unc, fig_m2_risk, fig_m2_educ, nrow=1)

ggsave("../LaTeX/_fig_models.pdf", plot=last_plot(), width = 24, height = 10, units = "cm")

```

################################################################################
##### EXPERIMENT 3 Straddle Tie

```{r}
# EXP3: vote shares of p(0) and p'(1) ;
table(exp_3$vote_unc)
prop.table(table(exp_3$vote_unc))
t.test(exp_3$vote_unc, mu=.5)
```

```{r}
# EXP3: vote shares in cases of complete overlap (v_i=p_A=p_B) ;
table(exp_3[exp_3$lr_i==exp_3$lr_A,]$vote)
prop.table(table(exp_3[exp_3$lr_i==exp_3$lr_A,]$vote))
t.test(exp_3[exp_3$lr_i==exp_3$lr_A,]$vote, mu=.5)
```

################################################################################
##### EXPERIMENT 4 REFLECTED TIE ;

```{r}
# EXP4: vote shares of p(0) and p' (1) ;
table(exp_4$vote_unc)
prop.table(table(exp_4$vote_unc))
t.test(exp_4$vote_unc, mu=.5)
```

################################################################################
##### EXPERIMENT 5 (PRECISE, RISKY, AND UNCERTAIN CHOICES)

# vignette 1 ("B=precise")
```{r}
# EXP51: vote shares of p_A (0) and p_B (1) ;
table(exp_5_1$vote)
prop.table(table(exp_5_1$vote))
t.test(exp_5_1$vote, mu=.5)
t.test(ifelse(exp_5_1$vote==0,1,0), mu=.5)
```

```{r}
# EXP51: bivariate model ;
m51 <- glm(vote ~ ut_B_A, family = binomial(link = "logit"), data=exp_5_1)
summary(m51)
gofm(m51)
```

```{r}
# EXP51: bivariate model (squared) ;
m51_sq <- glm(vote ~ ut_B_A_sq, family = binomial(link = "logit"), data=exp_5_1)
summary(m51_sq)
gofm(m51_sq)
```

################################################################################
# vignette 2 ("B=risky")
```{r}
# EXP52: vote shares of p_A(0) and p_B(1) ;
table(exp_5_2$vote)
prop.table(table(exp_5_2$vote))
t.test(exp_5_2$vote, mu=.5)
t.test(ifelse(exp_5_2$vote==0,1,0), mu=.5)
```

```{r}
# EXP52: vote shares of p(0) and p'(1)
table(exp_5_2$vote_unc)
prop.table(table(exp_5_2$vote_unc))
t.test(exp_5_2$vote_unc, mu=.5)
t.test(ifelse(exp_5_2$vote_unc==0,1,0), mu=.5)
```

```{r}
# EXP52: bivariate model ;
m52 <- glm(vote_unc ~ ut_unc_cer, family = binomial(link = "logit"), data=exp_5_2)
summary(m52)
gofm(m52)
```

```{r}
# EXP52: bivariate (squared) ;
m52_sq <- glm(vote_unc ~ ut_unc_cer_sq, family = binomial(link = "logit"), data=exp_5_2)
summary(m52_sq)
gofm(m52_sq)
```

```{r}
# EXP52: multivariate model with interaction ;
m52_int <- glm(vote_unc ~ ut_unc_cer * (lr_i_unc + risk + education), 
               family = binomial(link = "logit"), data=exp_5_2)
summary(m52_int)
gofm(m52_int)
```

```{r}
# EXP52: multivariate model with interaction (squared);
m52_int_sq <- glm(vote_unc ~ ut_unc_cer_sq * (lr_i_unc + risk + education), 
               family = binomial(link = "logit"), data=exp_5_2)
summary(m52_int_sq)
gofm(m52_int_sq)
```


################################################################################
# vignette 3 ("B=non-position")
```{r}
# EXP53: vote shares of p_A(0) and p_B(1) ;
table(exp_5_3$vote)
prop.table(table(exp_5_3$vote))
t.test(exp_5_3$vote, mu=.5)
t.test(ifelse(exp_5_3$vote==0,1,0), mu=.5)
```

```{r}
# EXP53: vote shares of p(0) and p'(1)
table(exp_5_3$vote_unc)
prop.table(table(exp_5_3$vote_unc))
t.test(exp_5_3$vote_unc, mu=.5)
t.test(ifelse(exp_5_3$vote_unc==0,1,0), mu=.5)
```

```{r}
# EXP53: bivariate model ;
m53 <- glm(vote_unc ~ ut_unc_cer, family = binomial(link = "logit"), data=exp_5_3)
summary(m53)
gofm(m53)
```

```{r}
# EXP53: bivariate model (squared);
m53_sq <- glm(vote_unc ~ ut_unc_cer_sq, family = binomial(link = "logit"), data=exp_5_3)
summary(m53_sq)
gofm(m53_sq)
```

```{r}
# EXP53: multivariate model with interaction ;
m53_int <- glm(vote_unc ~ ut_unc_cer * (lr_i_unc + risk + education), 
               family = binomial(link = "logit"), data=exp_5_3)
summary(m53_int)
gofm(m53_int)
```

################################################################################
# vignette 4 ("B=internal conflict")

```{r}
# EXP54: vote shares of p_A(0) and p_B(1) ;
table(exp_5_4$vote)
prop.table(table(exp_5_4$vote))
t.test(exp_5_4$vote, mu=.5)
t.test(ifelse(exp_5_4$vote==0,1,0), mu=.5)
```

```{r}
# EXP54: vote shares of p(0) and p'(1) ;
table(exp_5_4$vote_unc)
prop.table(table(exp_5_4$vote_unc))
t.test(exp_5_4$vote_unc, mu=.5)
t.test(ifelse(exp_5_4$vote_unc==0,1,0), mu=.5)
```

```{r}
# EXP54: bivariate model ;
m54 <- glm(vote_unc ~ ut_unc_cer, family = binomial(link = "logit"), data=exp_5_4)
summary(m54)
gofm(m54)
```

```{r}
# EXP54: bivariate model (squared);
m54_sq <- glm(vote_unc ~ ut_unc_cer_sq, family = binomial(link = "logit"), data=exp_5_4)
summary(m54_sq)
gofm(m54_sq)
```

```{r}
# EXP54: multivariate model with interaction ;
m54_int <- glm(vote_unc ~ ut_unc_cer * (lr_i_unc + risk + education), 
               family = binomial(link = "logit"), data=exp_5_3)
summary(m54_int)
gofm(m54_int)
```

```{r}
# EXP54: multivariate model with interaction (squared);
m54_int_sq <- glm(vote_unc ~ ut_unc_cer_sq * (lr_i_unc + risk + education), 
               family = binomial(link = "logit"), data=exp_5_3)
summary(m54_int_sq)
gofm(m54_int_sq)
```
```{r}
# EXP5: compare the four groups;

# no difference between choice under certainty and risk ;
t.test(exp_5_1$vote, exp_5_2$vote, mu=0)

# no difference between the choice-under-uncertainty scenarios ;
t.test(exp_5_3$vote, exp_5_4$vote, mu=0)

# differences between (1) choice under certainty or risk and (2) uncertainty

t.test(exp_5_1$vote, exp_5_3$vote, mu=0)
t.test(exp_5_1$vote, exp_5_4$vote, mu=0)

t.test(exp_5_2$vote, exp_5_3$vote, mu=0)
t.test(exp_5_2$vote, exp_5_4$vote, mu=0)

# joint test via logistic regression
m5 <- glm(vote ~ factor(vignette), data= exp_5, family=binomial(link = "logit"))
summary(m5)
confint(m5)

```

```{r}
screenreg(list(m1,m2,m51,m52,m53,m54), digits=3, dcolumn = TRUE, booktabs = TRUE, 
       include.bic=FALSE, use.packages = FALSE, single.row=FALSE, caption.above=TRUE, 
       fontsize="footnotesize", label = "TAB_APP_BI", 
       caption = "Discriminating Risk and Uncertainty (Abs. Distances)", 
       custom.coef.names = c("Intercept", "$u(p_B)-u(p_A)$","$u(p')-u(p)$"),
       custom.model.names=c("E1","E2","E5:1","E5:2","E5:3","E5:4"),
       custom.note="Notes: Coefficients are obtained from multilevel logit models based upon the forced-choice experiments E2-5.") 
```

```{r}
screenreg(list(m1_sq,m2_sq,m51_sq,m52_sq,m53_sq,m54_sq), digits=3, dcolumn = TRUE, 
      booktabs = TRUE, include.bic=FALSE, use.packages = FALSE, single.row=FALSE, 
      caption.above=TRUE, fontsize="footnotesize", label = "TAB_APP_BI_SQ", 
       caption = "Discriminating Risk and Ambiguity (Squared Distances)", 
       custom.coef.names = c("Intercept", "$u_{\text{sq}}(p_B)-u_{\text{sq}}(p_A)$","$u(p')-u(p)$"),
       custom.model.names=c("E1","E2","E5:1","E5:2","E5:3","E5:4"),
       custom.note="Notes: Coefficients are obtained multilevel logit models based on E2-5.") 
```


################################################################################
##### VOTING UNDER RISK VS AMBIGUITY ; 

# pooled models ;

```{r}
# EXP2-5: additive model ;
pooled_ra_1 <- glm(vote_unc ~ ut_unc_cer + amb, data=exp_ra,
                  family = binomial(link = "logit"))
summary(pooled_ra_1)
gofm(pooled_ra_1)
```

```{r}
# EXP2-5: interactive model  ;
pooled_ra_2 <- glm(vote_unc ~ ut_unc_cer * amb, data=exp_ra,
                  family = binomial(link = "logit"))
summary(pooled_ra_2)
```


# partially pooled models ;

```{r}
# EXP2-5: additive model  ;
ppooled_ra_1 <- glmer(vote_unc ~ ut_unc_cer + amb + (1|i), data=exp_ra,
                  family = binomial(link = "logit"))
summary(ppooled_ra_1)
```


```{r}
# EXP2-5: interactive model  ;
ppooled_ra_2 <- glmer(vote_unc ~ ut_unc_cer * amb + (1|i), data=exp_ra,
                  family = binomial(link = "logit"))
summary(ppooled_ra_2)
```

```{r}
screenreg(list(ppooled_ra_2), digits=3, dcolumn = TRUE, booktabs = TRUE, include.bic=FALSE, use.packages = FALSE, single.row=TRUE, caption.above=TRUE, 
       label = "TAB_PARTIAL", 
       caption = "Discriminating Risk and Ambiguity", 
       custom.coef.names = c("Intercept", "$u(p')-u(p)$", "ambiguity", "$[u(p')-u(p)]$ * ambiguity"),
       custom.model.names=c("Model 1"),
       custom.note="Notes: Coefficients are obtained from multilevel logit models based on E2-5.") 
```

```{r}
screenreg(list(pooled_ra_1, pooled_ra_2, ppooled_ra_1, ppooled_ra_2), digits=3, dcolumn = TRUE, booktabs = TRUE, include.bic=FALSE, use.packages = FALSE, single.row=FALSE, caption.above=TRUE, fontsize="footnotesize",
       label = "TAB_APP_RA", 
       caption = "Discriminating Risk and Ambiguity", 
       custom.coef.names = c("Intercept", "$u(p')-u(p)$", "ambiguity", "$[u(p')-u(p)]$ * ambiguity"),
       custom.model.names=c("pooled, add.", "pooled, int.", "multi., add.", "multi., int" ),
       custom.note="Notes: Coefficients are obtained from pooled and multilevel logit models based upon the forced-choice experiments E2-5.") 
```

```{r}
# EXP2-5: marginal effects plot ;
amb0 <- as.numeric(predict(pooled_ra_2, type="response", newdata=data.frame(ut_unc_cer=c(0,0), amb = c(0,1)))[1])
amb1 <- as.numeric(predict(pooled_ra_2, type="response", newdata=data.frame(ut_unc_cer=c(0,0), amb = c(0,1)))[2])

ggpredict(pooled_ra_1, c("ut_unc_cer", "amb")) %>% 
  plot(color="bw", add.data=TRUE, jitter=0.02) + 
  geom_segment(aes(x = 0, y = amb0, xend = 0, yend = amb1), linewidth=2) +
  geom_segment(aes(x = 0, y = amb0, xend = -8, yend = amb0), linetype="dotted") +
  geom_segment(aes(x = 0, y = amb1, xend = -8, yend = amb1), linetype="dotted") +
  xlab("utility difference u(c[p'])-u(p)") + 
  ylab("vote probability [Pr(v=p')]") +
  ggtitle("") + theme_bw() + theme(legend.position="bottom", legend.title=element_blank())

ggsave("../LaTeX/_fig_partial.pdf", plot=last_plot(), width = 12, height = 12, units = "cm")
```
